---
title: "Tangye_InbornErrorsImmunity"
output: html_document
date: "2023-11-20"
---

```{r setup, include=FALSE}
library(tidyverse)
library(hgnc)
```

## Open hgnc
```{r load hgnc ref}
#check version, update
hgnc::last_update()
(url <- latest_archive_url())
hgnc_dataset <- import_hgnc_dataset(url)
hgnc_dataset<-hgnc_dataset %>% tidyr::unnest(., cols = alias_symbol)
hgnc_dataset<-hgnc_dataset %>% tidyr::unnest(., cols = prev_symbol)

hgnc_gene<-hgnc_dataset %>%
  select(hgnc_id,symbol)

hgnc_alias<-hgnc_dataset %>%
  select(hgnc_id,alias_symbol)

hgnc_prev<-hgnc_dataset %>%
  select(hgnc_id,prev_symbol)

```

Read in gene list data:

Tangye SG, Al-Herz W, Bousfiha A, Cunningham-Rundles C, Franco JL, Holland SM, Klein C, Morio T, Oksenhendler E, Picard C, Puel A, Puck J, Seppänen MRJ, Somech R, Su HC, Sullivan KE, Torgerson TR, Meyts I. Human Inborn Errors of Immunity: 2022 Update on the Classification from the International Union of Immunological Societies Expert Committee. J Clin Immunol. 2022 Oct;42(7):1473-1507. doi: 10.1007/s10875-022-01289-3. Epub 2022 Jun 24. PMID: 35748970; PMCID: PMC9244088.

```{r}
Tangye_2023 <- read_excel("~/Documents/NIH_reading/Gene List/10875_2022_1289_MOESM2_ESM.xlsx")
```

Join Tangye_2023 to HGNC official gene symbol table for 1:1 matching, no clean up required
```{r}
#Join supp table to HGNC official for 1:1 matching
Tangye_2023_hgnc <-Tangye_2023 %>%
  left_join(.,hgnc_gene, by = c("Genetic defect" = "symbol")) %>%
  distinct()
```

Identify number of rows requiring cleanup
```{r}
sum(is.na(Tangye_2023_hgnc$hgnc_id))
```

```{r}
#identify which rows did not match
missing_IEUS<-which(is.na(Tangye_2023_hgnc$hgnc_id))
missing_IEUS_list<-Tangye_2023_hgnc[missing_IEUS,"Genetic defect"]
missing_IEUS_list
write.csv(missing_IEUS_list, "missing_IEUS_list.csv")

```

